---
title: "District 2 squirrel model"
author: Lorna Haworth
output: 
---
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

```{r load libraries} 
knitr::opts_knit$set(root.dir = "C:/R_studio/Research/Median-Barriers")
getwd()

library(tidyverse)
library(conflicted)
library(lmtest)     # for lrtest()
library(bbmle)      # for ICtab()
library(stringr)
library(emmeans)
library(multcomp)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("summarize", "dplyr")
conflict_prefer("case_where", "dplyr")

```
Read in CSV files that have data we collected, HPMS, and NCLD distance to data.
```{r read in data}

#District 2 random points
D2_random_sites <- read.csv("d2_random_sites_HPMS_updatedLEH.csv")

#subset, select random rows into new data frame


#District 2 western grey squirrel
D2_squirrel_hits <- read.csv("d2_squirrel_hits_HPMS_updatedLEH.csv")

```
Make the data frames look better by only using necessary data. Including median data we collected, HPMS data, and NCLD distance to data.

Add date information - can't add date information for roadkill data frame bc then can't join together
```{r make new data frames with simplified information from csv }
D2_random <- subset(D2_random_sites, select=c(MedianType,
                                               AADT,
                                               bldg_dist,
                                               nhd_dist,
                                               ow_dist,
                                               dev_dist,
                                               dev_int_dist,
                                               barren_dist,
                                               forest_dist,
                                               shrub_dist,
                                               grassland_dist,
                                              cultivated_dist,
                                              wetland_dist,
                                               wet_dist))
                                              
#add columns to random points dataframe for to distinguish random and animal condition
D2_random$animal<-"random" 
D2_random$condition <- "none"


D2_squirrel <- subset(D2_squirrel_hits, select = c(animal, 
                                           condition, 
                                           MedianType,
                                           AADT,
                                           bldg_dist,
                                           nhd_dist,
                                           ow_dist,
                                           dev_dist,
                                           dev_int_dist,
                                           barren_dist,
                                           forest_dist,
                                           shrub_dist,
                                           grassland_dist,
                                           cultivated_dist,
                                           wetland_dist,
                                           wet_dist))

```


```{r organize and group data}

#make combined data frame 
squir_rand <- rbind(D2_squirrel, D2_random) 


#keep only one median type in structure column, new data frame with single classes
median_deer <- median_ug %>%
  as.data.frame() %>%
  mutate(MedianType = case_when(
    str_detect(MedianType, "thrie beam") ~ "thrie beam", # If "thrie beam" is anywhere in the string, change to "thrie beam"
    str_detect(MedianType, "concrete") ~ "concrete",
    str_detect(MedianType, "cable") ~ "cable",
    str_detect(MedianType, "gravel") & str_detect(MedianType, "vegetative") ~ "vegetative", # If both "gravel" and "vegetative" are present, change to "vegetative"
    TRUE ~ MedianType # Keep other values as they are
  ))

#Create binary indicator for structures
median_deer$structure <- 0 # Create structure column and assign it a default value of 0
median_deer2 <- median_deer %>%
  mutate(structure = case_when(
    str_detect(MedianType, "thrie beam") ~ 1, # If MedianType contains "thrie beam", change structure to 1
    str_detect(MedianType, "concrete") ~ 1, # If MedianType contains "concrete", change structure to 1
    str_detect(MedianType, "cable") ~ 1, # If MedianType contains "cable", change structure to 1
    TRUE ~ structure # Keep other values as they are
  ))

#Create binary indicator for WVC
median_deer2$hit <- 0 # Create hit column and assign it a default value of 0
median_deer3 <- median_deer2 %>%
  mutate(hit = case_when(
    str_detect(condition, "Dead") ~ 1,
    str_detect(condition, "Injured") ~ 1,
    TRUE ~ hit))



```


```{r glm model with  binomial distribution}
#model with neg binom
m1 <- glm(hit ~ -1 + as.factor(MedianType),
                     data = median_deer3,
                      family = binomial(link="logit"))
plogis(coef(m1))

#null
m0 <- glm(hit ~ 1,
          data = median_deer3,
          family=binomial(link="logit"))

#adding some predictors...
m2 <- glm(hit ~ MedianType + nhd_dist + bldg_dist ,
                     data = median_deer3,
                      family = binomial(link="logit"))
lrtest(m1, m2)
Anova(m2)

#all NCLD distance to calcs
m3 <- glm(hit ~ 0 + MedianType + bldg_dist + nhd_dist + ow_dist + dev_dist + dev_int_dist + barren_dist + forest_dist + shrub_dist + grassland_dist + cultivated_dist + wetland_dist + wet_dist,
                     data = median_deer3,
                      family = binomial(link="logit"))
coef(m3)
Anova(m3)
lrtest(m2, m3)


# looking for best way to see which median types are better/worse
results <- emmeans(m1, pairwise ~ MedianType, type = "response", adjust = "tukey") #pairwise comparison of each median type to see which one has strongest effect
results$emmeans
results$contrasts
contrast(emmeans(m1, ~ MedianType, type = "response", method = "pairwise"))


#box plot to visually show results from above


ggplot(median_deer3, aes(MedianType, exp(predict(m1)))) + geom_point()

# trying tukey test
summary(glht(m1, mcp(MedianType = "Tukey")))
```

random effects brainstorm: username from roadkill data
```{r mixing it up}
###GLMM goes here
```


