---
title: "Practicing the model deer"
author: Lorna Haworth
output: 
---
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

BRAINSTORMING NOTES
-data frame has response variable (roadkill) and possible predictors (NCLD, NHD, building)
-no empty values in rows

deer first, then animals in order of size (large-small)
maybe mtn lion next
sin wave on residual chart ?? why that shape

```{r load libraries} 
knitr::opts_knit$set(root.dir = "C:/R_studio/Research/Median-Barriers")
getwd()

library(tidyverse)
library(conflicted)
library(lmtest)     # for lrtest()
library(bbmle)      # for ICtab()
library(glmmTMB)    #to run simulation
library(DHARMa)     #to run simulation
library(stringr)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("summarize", "dplyr")

```

```{r read in data}

#District 2 random points
D2_random_sites <- read.csv("d2_random_sites_HPMS.csv")

#District 2 mule deer
D2_deer_hits <- read.csv("d2_deer_hits_HPMS.csv")

```


Make the data frames look better by only using necessary data
```{r make new data frames with simplified information from csv }

D2_random <- subset(D2_random_sites, select=c(#cid,
                                               MedianType,
                                               LANE_WIDTH,
                                               MEDIAN_TYP,
                                               MEDIAN_WID))
                                              # SecondaryAttribute, 
                                              # MedianWidth, 
                                              # RoadsideBarrier))

D2_deer <- subset(D2_deer_hits, select = c(animal, 
                                           condition, 
                                           MedianType,
                                           LANE_WIDTH,
                                           MEDIAN_TYP,
                                           MEDIAN_WID))

#add columns to random points dataframe for to distinguish random and animal condition
D2_random$animal<-"random" 
D2_random$condition <- "none"
```
LORNA START HERE

```{r organize and group data}

#make combined data frame 

#squirrels
point_median_ug <- rbind(D2_random, D2_squirrel) #use in binomial model

point_median_sq <- point_median_ug %>%  #use in poisson
  group_by(MedianType, condition) %>%
  summarize(count=n(), .groups = "drop")

#keep only one median type in structure column, new data frame with single classes
point_median_sq2 <- point_median_sq %>%
  as.data.frame() %>%
  mutate(MedianType = case_when(
    str_detect(MedianType, "thrie beam") ~ "thrie beam", # If "thrie beam" is anywhere in the string, change to "thrie beam"
    str_detect(MedianType, "concrete") ~ "concrete",
    str_detect(MedianType, "cable") ~ "cable",
    str_detect(MedianType, "gravel") & str_detect(MedianType, "vegetative") ~ "vegetative", # If both "gravel" and "vegetative" are present, change to "vegetative"
    TRUE ~ MedianType # Keep other values as they are
  ))

#create binary indicator for structures, add opaque (thrie + concrete) vs not (cable)
point_median_sq2$structure <- 0 # Create structure column and assign it a default value of 0
point_median_sq2 <- point_median_sq2 %>%
  mutate(structure = case_when(
    str_detect(MedianType, "thrie beam") ~ 1, # If MedianType contains "thrie beam", change structure to 1
    str_detect(MedianType, "concrete") ~ 1, # If MedianType contains "concrete", change structure to 1
    str_detect(MedianType, "cable") ~ 1, # If MedianType contains "cable", change structure to 1
    TRUE ~ structure # Keep other values as they are
  ))


#deer
deer_median_ug <- rbind(D2_random, D2_deer)

deer_median <- deer_median_ug %>%  
  group_by(MedianType, condition) %>%
  summarize(count=n(), .groups = "drop")

deer_median_2 <- deer_median %>%
  as.data.frame() %>%
  mutate(MedianType = case_when(
    str_detect(MedianType, "thrie beam") ~ "thrie beam", # If "thrie beam" is anywhere in the string, change to "thrie beam"
    str_detect(MedianType, "concrete") ~ "concrete",
    str_detect(MedianType, "cable") ~ "cable",
    str_detect(MedianType, "gravel") & str_detect(MedianType, "vegetative") ~ "vegetative", # If both "gravel" and "vegetative" are present, change to "vegetative"
    TRUE ~ MedianType # Keep other values as they are
  ))

deer_median_2$structure <- 0 # Create structure column and assign it a default value of 0
deer_median_2 <- deer_median_2 %>%
  mutate(structure = case_when(
    str_detect(MedianType, "thrie beam") ~ 1, # If MedianType contains "thrie beam", change structure to 1
    str_detect(MedianType, "concrete") ~ 1, # If MedianType contains "concrete", change structure to 1
    str_detect(MedianType, "cable") ~ 1, # If MedianType contains "cable", change structure to 1
    TRUE ~ structure # Keep other values as they are
  ))


```

TO DO make binomial model with same predictors 
Should I use means parameterization in the model? Is roadkill categorical data?
```{r glm model}
#poisson family
m1_D2_squirrel <-glm(count ~ MedianType + condition,
                     data = point_median_sq2,
                     family = poisson(link = "log"))
summary(m1_D2_squirrel)

m1_deer <- glm(count ~ 1,
               data = deer_median_2,
               family = poisson(link="log"))

m2_D2_deer <-glm(count ~ MedianType + condition,
                     data = deer_median_2,
                     family = poisson(link = "log"))
lrtest(m1_deer, m2_D2_deer) #m2 is better than the null

#binomial family
m2_D2_squirrel <-glm(count ~ MedianType + condition,  #i think make condition binomial and use that as response ?
                     data = point_median_sq2,
                     family = binomial(link = "logit"))
summary(m2_D2_squirrel)

```


explain what this is doing, is the model good or bad for the data? (bad)
```{r test model against simulated points}
#with model 1 squirrel (Poisson)
simulation_output_1 <- simulateResiduals(fittedModel = m1_D2_squirrel)
plot(simulation_output_1)
testUniformity(simulation_output) #why the sin wave shape? does that indicate something with the data?

#with model 1 deer (poisson)
simulation_output_2 <- simulateResiduals(fittedModel = m2_D2_deer)
plot(simulation_output_2)
testUniformity(simulation_output_2)

#null of deer just to see results
simulation_output_3 <- simulateResiduals(fittedModel = m1_deer)
plot(simulation_output_3)
testUniformity(simulation_output_3)

#with model 2 (binomial)
simulation_output2 <- simulateResiduals(fittedModel = m2_D2_squirrel)
plot(simulation_output2)
testUniformity(simulation_output2)


```


