---
title: "Practice with datasets"
author: Lorna Haworth
output: 
---
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
First, set working directory and load appropriate libraries.

BRAINSTORMING NOTES
data frame has response variable (roadkill) and possible predictors (NCLD, NHD, building)
no empty values in rows

what is the median type telling you in a random points? how to include random 
 --> are they correlated

how to make a model with only presence data (no absence)

```{r load libraries} 
knitr::opts_knit$set(root.dir = "C:/R_studio/Research/Median-Barriers")
getwd()

library(tidyverse)
library(conflicted)
library(glmmTMB)
library(DHARMa)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("summarize", "dplyr")
?group_by
```

```{r read in data}
District2_squirrels <- read.csv("D2_squirrel_hits.csv")
District2_random <- read.csv("D2_random_sites.csv")
```


Make the data frames look beter by only using necessary data
```{r make new data frames with simplified information from csv }
D2_squirrel_road <- subset(District2_squirrels, select=c(#nid,
                                                         animal, 
                                                          condition, 
                                                          #created_da, 
                                                          MedianType)) 
                                                          #SecondaryAttribute, 
                                                          #MedianWidth, 
                                                          #RoadsideBarrier))
D2_random <- subset(District2_random, select=c(#cid,
                                               MedianType))
                                              # SecondaryAttribute, 
                                              # MedianWidth, 
                                              # RoadsideBarrier))

#D2.random <- mutate(as.factor[animal="random"]) #HELP need to add column to random points to show no roadkill
D2_random$animal<-"random" 
D2_random$condition <- "none"
```


TO DO: make table that shows totals for number of roadkill for every median type compared against if it was a roadkill point or simulated point
```{r organize data}
#collapse(num.RK ~ MedianType )
#make combined data frame called "D2.squirrel.com"

point_median_ug <- rbind(D2_random, D2_squirrel_road)

point_median <- point_median_ug %>% 
  group_by(MedianType, condition) %>%
  summarize(count=n(), .groups = "drop")



```

TO DO
```{r glm model}
m1_D2_squirrel <-glm(count ~ MedianType + condition,
                     data = point_median,
                     family = poisson(link = "log"))
summary(m1_D2_squirrel)

```
```{r,}
simulation_output <- simulateResiduals(fittedModel = m1_D2_squirrel)
plot(simulation_output)
testUniformity(simulation_output)

```


