---
title: "Practice model"
author: Lorna Haworth
output: 
---
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.
First, set working directory and load appropriate libraries.

BRAINSTORMING NOTES
data frame has response variable (roadkill) and possible predictors (NCLD, NHD, building)
no empty values in rows

what is the median type telling you in a random points? how to include random 
 --> are they correlated

how to make a model with only presence data (no absence)

```{r load libraries} 
knitr::opts_knit$set(root.dir = "C:/R_studio/Research/Median-Barriers")
getwd()

library(tidyverse)
library(conflicted)
library(glmmTMB) #to run simulation
library(DHARMa) #to run simulation
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("summarize", "dplyr")

```

```{r read in data}
District2_squirrels <- read.csv("D2_squirrel_hits.csv")
District2_random <- read.csv("D2_random_sites.csv")
```


Make the data frames look beter by only using necessary data
```{r make new data frames with simplified information from csv }
D2_squirrel_road <- subset(District2_squirrels, select=c(#nid,
                                                         animal, 
                                                          condition, 
                                                          #created_da, 
                                                          MedianType)) 
                                                          #SecondaryAttribute, 
                                                          #MedianWidth, 
                                                          #RoadsideBarrier))
D2_random <- subset(District2_random, select=c(#cid,
                                               MedianType))
                                              # SecondaryAttribute, 
                                              # MedianWidth, 
                                              # RoadsideBarrier))


D2_random$animal<-"random" 
D2_random$condition <- "none"
```

separate median types:
library(stringr)
deer_comb <- deer_comb %>%
  as.data.frame() %>%
  mutate(MedianType = case_when(
    str_detect(MedianType, "thrie beam") ~ "thrie beam", # If "thrie beam" is anywhere in the string, change to "thrie beam"
    str_detect(MedianType, "concrete") ~ "concrete", 
    str_detect(MedianType, "cable") ~ "cable",
    str_detect(MedianType, "gravel") & str_detect(MedianType, "vegetative") ~ "vegetative", # If both "gravel" and "vegetative" are present, change to "vegetative"
    TRUE ~ MedianType # Keep other values as they are
  ))


```{r organize data}
#collapse(num.RK ~ MedianType )
#make combined data frame called "D2.squirrel.com"

point_median_ug <- rbind(D2_random, D2_squirrel_road) #use binomial model

point_median <- point_median_ug %>%  #use poisson
  group_by(MedianType, condition) %>%
  summarize(count=n(), .groups = "drop")



```

TO DO make binomial model with same predictors 
```{r glm model}
m1_D2_squirrel <-glm(count ~ MedianType + condition,
                     data = point_median,
                     family = poisson(link = "log"))
summary(m1_D2_squirrel)

```
```{r test model against simulated points}
simulation_output <- simulateResiduals(fittedModel = m1_D2_squirrel)
plot(simulation_output)
testUniformity(simulation_output)

```


